---
title: "Data_Analysis"
author: "Szu-Tung Chen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

## Read Cleaned File

```{r}
set.seed(1450)
library(MASS)
library(dplyr)
library(nnet)
library(ggplot2)
USA = read.csv("USA.csv")

White = subset(USA, Q290 == 840001)
```

## data processing

```{r}
df = White
df = subset(df, Q260=="Female")
df = subset(df, select = -c(Q261, Q263, Q46P, Q269, Q271, Q272, Q264, Q265, Q260, Q290, Q266, Q267, Q268))

hist(df$Q49)
abline(v=mean(df$Q49), col="red")

factorize = function(df_temp){
  df_temp[] = lapply(df_temp, function(x) if(is.character(x)) factor(x) else x)
  return(df_temp)
}


continue = TRUE
while (continue){
  n = nrow(df)
  train_indices = sample(1:n, 160)
  train_df = df[train_indices, ]
  test_df = df[-train_indices, ]
  train_df = factorize(train_df)
  test_df = factorize(test_df)
  
  check = 0
  for (col_name in colnames(df)) {
    if (is.factor(df[[col_name]])) {
      if (levels(test_df[[col_name]]) %in% levels(train_df[[col_name]])){
        check = 1 # need to resample
      }
    }
  }
  if (check == 0){continue = FALSE}
}
```

### Fit Full Model & Check Assumption

```{r}
full_model = lm(Q49 ~ ., train_df)
summary(full_model)
# Normality Assumption
qqnorm(full_model$residuals, ylab="Residuals")
qqline(full_model$residual)
# Check Mean Model and Constant Variance
plot(full_model$fitted, full_model$residual, xlab="Fitted",ylab="Residuals")
abline(h=0)
plot(full_model$fitted, abs(full_model$residual), xlab="Fitted",ylab="|Residuals|")
abline(h=0)

```
# Unusual Points
```{r}
# Leverage Point
hatv = hatvalues(full_model)
library(faraway)
halfnorm(hatv, nlab = 0, ylabs = "Leverages")
which(hatv>0.99) 
# Cook's distance
cook <- cooks.distance(full_model)
halfnorm(cook, nlab = 0, ylab="Cook’s distance")
cook_update = lm(Q49 ~ ., train_df, subset=(cook<0.05))
train_df = subset(train_df, cook<0.05)
summary(cook_update)
full_model = cook_update

shapiro.test(full_model$residual)
```



# Model Selection

```{r, message=FALSE, warning=FALSE}
# forward selection
library(olsrr)
ols_step_backward_p(full_model, prem=0.05, details=FALSE)
ols_step_forward_p(full_model, pent=0.05, details=FALSE)
ols_step_both_p(full_model, pent=0.05, prem=0.05, details=FALSE)
```


### check interaction terms

```{r}
check_interaction_significance <- function(model) {
  # Get the summary of the model
  summary_model <- summary(model)
  
  # Extract p-values
  p_values <- coef(summary_model)[, "Pr(>|t|)"]
  
  # Set the significance level
  significance_level <- 0.0001
  
  # Find interaction terms and check if any are significant
  interaction_terms <- grep(":", names(p_values))
  any(p_values[interaction_terms] < significance_level)
}

check_list = c("Q49", "Q266", "Q267", "Q268", "Q272")
total_variable_names = names(train_df)
selected_variable_names = c("Q285", "Q288", "Q286", "Q282")
significant_pairs = list()

# Remove variables in check_list from variable_names
variable_names = setdiff(total_variable_names, check_list)

for (i in 1:(length(variable_names) - 1)) {
  for (j in (i + 1):length(variable_names)) {
    if (j<=i) { break }
    var1 = variable_names[i]
    var2 = variable_names[j]

    mod = lm(Q49 ~ get(var1) * get(var2), data = train_df)
    if (check_interaction_significance(mod)) {
      cat("var1 =", var1, "var2 =", var2, "\n")
      significant_pairs[[length(significant_pairs) + 1]] = paste(var1, var2, sep=":")
    }
  }
}

new_model_formula = paste("Q49 ~ . +", paste(significant_pairs, collapse=" + "))
new_model = lm(as.formula(new_model_formula), data = train_df)
summary = summary(new_model)
summary

```



## Aggregate Q275
```{r}
levels(train_df$Q275)
levels(train_df$Q275) = c("UG", "Grad", "Grad", "ISCED4", "ISCED5", "ISCED3")

model275 = lm(as.formula(new_model_formula), data = train_df)
anova(new_model, model275)
summary(model275)$r.square
summary(model275)$adj.r.square
new_model = model275

```

Aggregated Model is Acceptable

## Aggregate Q276
```{r}
levels(train_df$Q276)

levels(train_df$Q276) = c("UG", "Grad", "Grad", "ISCED4", "ISCED5", "ISCED3")
model276 = lm(as.formula(new_model_formula), data = train_df)
anova(new_model, model276)
summary(model276)$r.square
summary(model276)$adj.r.square
new_model = model276
```
Aggregated model is acceptable


## Aggrefate Q278

```{r}
levels(train_df$Q278)
levels(train_df$Q278) = c("UG", "Grad", "secondary", "Grad", "ISCED45", "ISCED45", "secondary")
model278 = lm(as.formula(new_model_formula), data = train_df)
anova(new_model, model278)
summary(model278)$r.square
summary(model278)$adj.r.square
new_model = model278
```

Aggregated Model is Acceptable



```{r}
# without Q284
reduced_model = lm(Q49 ~ Q275 + Q276 + Q278 + Q285 + Q286 + 
    Q288 + Q275:Q278 + Q276:Q278, train_df)
anova(reduced_model, new_model)
summary(reduced_model)$r.square
summary(reduced_model)$adj.r.square
summary(new_model)$r.square
summary(new_model)$adj.r.square

```
## Aggregate Levels for Testing Data
```{r}
# transform levels in test
test_df = subset(test_df, test_df$Q281 != "JP,KG,TJ: Other")
levels(test_df$Q275) = c("UG", "Grad", "ISCED4", "ISCED5", "ISCED3")
levels(test_df$Q276) = c("UG", "Grad", "ISCED3", "Grad", "ISCED4", "ISCED5")
levels(test_df$Q278) = c("UG", "Grad", "secondary", "Grad", "ISCED45", "ISCED45", "secondary")
```
## Model Validation: Model Without Selection
```{r }
# RMSE function
rmse <- function(x,y){
  sqrt(mean((x-y)^2))
}
# Model Without Selection
pred_new = as.data.frame(predict(new_model, newdata = test_df))
pred_new[pred_new<=0] = 0
pred_new[pred_new>=10] = 10

predict(new_model, newdata = test_df)

rmse(new_model$fitted.values, train_df$Q49)
rmse(pred_new[,], test_df$Q49)
plot(test_df$Q49, pred_new[,], ylab="Predicted Life Satisfaction", xlab="Recorded Life Satisfaction")
abline(0,1, col="red")
abline(1.5,1, col="blue")
abline(-1.5,1, col="blue")

```

## Model Validation: Model With Selection
```{r warning=FALSE}
# Model With Selection
pred_red = as.data.frame(predict(reduced_model, newdata = test_df))
pred_red[pred_red<=0] = 0
pred_red[pred_red>=10] = 10

rmse(reduced_model$fitted.values, train_df$Q49)
rmse(pred_red[,], test_df$Q49)
plot(test_df$Q49, pred_red[,], ylab="Predicted Life Satisfaction", xlab="Recorded Life Satisfaction")
abline(0,1, col="red")
abline(1.5,1, col="blue")
abline(-1.5,1, col="blue")
```



## plot interactions and main effects
```{r}
plot_df = train_df
plot_df$y=new_model$fitted.values


which(coef(summary(reduced_model))[, "Pr(>|t|)"] < 0.05)

# interaction term Q275 Q278
ggplot(plot_df, aes(x=Q275, y=y, fill=Q278)) + geom_boxplot() + scale_x_discrete(guide = guide_axis(angle = 40)) + facet_wrap(Q278 ~ .)

# interaction term Q276 Q278
ggplot(plot_df, aes(x=Q276, y=y, fill=Q278)) + geom_boxplot() + scale_x_discrete(guide = guide_axis(angle = 40)) + facet_wrap(Q278 ~ .)



```



