---
title: "Data_Analysis"
author: "Szu-Tung Chen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1)
```

## Read Cleaned File

```{r message=FALSE}
set.seed(1450)
library(MASS)
library(dplyr)
library(nnet)
library(ggplot2)
library(quantreg)
USA = read.csv("USA.csv")

White = subset(USA, Q290 == 840001)
```

## data processing

```{r}
df = White
df = subset(df, Q260=="Male")

df = subset(df, select = -c(Q261, Q263, Q46P, Q269, Q271, Q272, Q264, Q265, Q260, Q290, Q266, Q267, Q268,Q289))

hist(df$Q49)
abline(v=mean(df$Q49), col="red")

factorize = function(df_temp){
  df_temp[] = lapply(df_temp, function(x) if(is.character(x)) factor(x) else x)
  return(df_temp)
}

continue = TRUE
while (continue){
  n = nrow(df)
  train_indices = sample(1:n, 220)
  train_df = df[train_indices, ]
  test_df = df[-train_indices, ]
  train_df = factorize(train_df)
  test_df = factorize(test_df)
  
  check = 0
  for (col_name in colnames(df)) {
    if (is.factor(df[[col_name]])) {
      if (levels(test_df[[col_name]]) %in% levels(train_df[[col_name]])){
        check = 1 # need to resample
      }
    }
  }
  if (check == 0){continue = FALSE}
}
summary(train_df)
```

### Fit Full Model & Check Assumption

```{r}
full_model = lm(Q49 ~ ., data=train_df)

# Normality Assumption
qqnorm(full_model$residuals, ylab="Residuals")
qqline(full_model$residual)
# Check Mean Model and Constant Variance
plot(full_model$fitted, full_model$residual, xlab="Fitted",ylab="Residuals")
abline(h=0)
plot(full_model$fitted, abs(full_model$residual), xlab="Fitted",ylab="|Residuals|")
abline(h=0)

shapiro.test(full_model$residual)

```
# Unusual Points
```{r}
# Leverage Point
hatv = hatvalues(full_model)
library(faraway)
halfnorm(hatv, nlab = 0, ylabs = "Leverages")
which(hatv>0.99) 
# Cook's distance
cook <- cooks.distance(full_model)
halfnorm(cook, nlab = 0, ylab="Cook’s distance")
cook_update = lm(Q49 ~ ., train_df, subset=(cook<0.025))
df = subset(train_df, cook<0.025)
summary(cook_update)
full_model = cook_update

shapiro.test(full_model$residual)
# Normality Assumption
qqnorm(full_model$residuals, ylab="Residuals")
qqline(full_model$residual)
# Check Mean Model and Constant Variance
plot(full_model$fitted, full_model$residual, xlab="Fitted",ylab="Residuals")
abline(h=0)
plot(full_model$fitted, abs(full_model$residual), xlab="Fitted",ylab="|Residuals|")
abline(h=0)
```



# Model Selection

```{r, message=FALSE, warning=FALSE}

# # Check Colinearity
# X = model.matrix(full_model)[, -1]
# which(vif(X) >10)
# summary(full_model)
# forward selection
library(olsrr)
ols_step_backward_p(full_model, prem=0.05, details=FALSE)
ols_step_forward_p(full_model, pent=0.05, details=FALSE)
ols_step_both_p(full_model, pent=0.05, prem=0.05, details=FALSE)
# adjusted R square
# library(leaps)
# b = regsubsets(df$Q49 ~ ., df, really.big=T)
# rs = summary(b)
# which.max(rs$adjr2) - 1
```


### check interaction terms

```{r}
check_interaction_significance <- function(model) {
  # Get the summary of the model
  summary_model <- summary(model)
  
  # Extract p-values
  p_values <- coef(summary_model)[, "Pr(>|t|)"]
  
  # Set the significance level
  significance_level <- 0.0001
  
  # Find interaction terms and check if any are significant
  interaction_terms <- grep(":", names(p_values))
  any(p_values[interaction_terms] < significance_level)
}

check_list = c("Q49", "Q266", "Q267", "Q268", "Q272")
total_variable_names = names(train_df)
selected_variable_names = c("Q285", "Q288", "Q286", "Q282")
significant_pairs = list()

# Remove variables in check_list from variable_names
variable_names = setdiff(total_variable_names, check_list)

for (i in 1:(length(variable_names) - 1)) {
  for (j in (i + 1):length(variable_names)) {
    if (j<=i) { break }
    var1 = variable_names[i]
    var2 = variable_names[j]

    mod = lm(Q49 ~ get(var1) * get(var2), data = train_df)
    if (check_interaction_significance(mod)) {
      cat("var1 =", var1, "var2 =", var2, "\n")
      significant_pairs[[length(significant_pairs) + 1]] = paste(var1, var2, sep=":")
    }
  }
}

new_model_formula = paste("Q49 ~ . +", paste(significant_pairs, collapse=" + "))
new_model = lm(as.formula(new_model_formula), data = train_df)
summary = summary(new_model)
summary

```



## Aggregate Q275
```{r}
levels(train_df$Q275)
levels(train_df$Q275) = c("UG", "Master", "secondary", "Doctoral", "ISCED4", "ISCED5", "secondary")

model275 = lm(as.formula(new_model_formula), data = train_df)
anova(new_model, model275)
summary(model275)$r.square
summary(model275)$adj.r.square
new_model = model275
```

Aggregated Model is Acceptable

## Aggregate Q281
```{r}
levels(train_df$Q281)
levels(train_df$Q281) = c("Clerical",
                          "Other",
                          "Higher administrative",
                          "Other",
                          "Professional and technical",
                          "Sales",
                          "Other",
                          "Service",
                          "Skilled worker",
                          "Other")
model281 = lm(as.formula(new_model_formula), data = train_df)
anova(new_model, model281)
summary(model281)$r.square
summary(model281)$adj.r.square
new_model = model281
```

Aggregated model is acceptable.


## Aggrefate Q276

```{r}
levels(train_df$Q276)
levels(train_df$Q276) = c("UG", "Grad", "secondary", "Grad", "ISCED45", "ISCED45", "secondary")
model276 = lm(as.formula(new_model_formula), data = train_df)
anova(new_model, model276)
summary(model276)$r.square
summary(model276)$adj.r.square
new_model = model276
```

Aggregated Model is Acceptable

# Aggregate Q282

```{r}
levels(train_df$Q282)
levels(train_df$Q282) = c("Clerical",
                          "Higher administrative",
                          "Other",
                          "Professional and technical",
                          "Sales",
                          "Other",
                          "Service",
                          "Other",
                          "Other")
model282 = lm(as.formula(new_model_formula), data = train_df)
anova(new_model, model282)
summary(model282)$r.square
summary(model282)$adj.r.square
new_model = model282
```
Aggregation is acceptable

# Aggregate Q286
```{r}
temp_df = train_df
levels(temp_df$Q286)
levels(temp_df$Q286) = c("Just get by",
                          "Save money",
                          "borrow",
                          "borrow")
model286 = lm(as.formula(new_model_formula), data = temp_df)
anova(new_model, model286)
summary(model286)$r.square
summary(model286)$adj.r.square
```



```{r}

# Test-based selection
reduced_model = lm(Q49 ~ Q275 + Q276+ Q281+
    Q282 + Q286 + Q288 + Q275:Q281 + Q276:Q282 + Q282:Q286, train_df)
anova(reduced_model, new_model)
summary(reduced_model)$r.square
summary(reduced_model)$adj.r.square
summary(new_model)$r.square
summary(new_model)$adj.r.square
```
## Aggregate Levels for Testing Data
```{r}
levels(test_df$Q275) = c("UG", "Doctoral", "Master", "ISCED4", "ISCED5", "secondary")
levels(test_df$Q276) = c("UG", "Grad", "Grad", "ISCED45", "ISCED45", "secondary")
levels(test_df$Q281) = c("Clerical", "Higher administrative", "Other", "Professional and technical", "Sales","Other", "Service", "Skilled worker", "Other")
levels(test_df$Q282) = c("Clerical", "Higher administrative", "Professional and technical", "Sales","Other", "Service", "Other")
```
## Model Validation: Model Without Selection
```{r warning=FALSE}
# RMSE function
rmse <- function(x,y){
  print(y)
  sqrt(mean((x-y)^2))
}
# Model Without Selection
pred_new = as.data.frame(predict(new_model, newdata = test_df))
pred_new[pred_new<=0] = 0
pred_new[pred_new>=10] = 10

rmse(new_model$fitted.values, train_df$Q49)
rmse(pred_new[,], test_df$Q49)
plot(test_df$Q49, pred_new[,], ylab="Predicted Life Satisfaction", xlab="Recorded Life Satisfaction")
abline(0,1, col="red")
abline(1.5,1, col="blue")
abline(-1.5,1, col="blue")
```

## Model with selection
```{r}
pred_red = as.data.frame(predict(reduced_model, newdata = test_df))
pred_red[pred_red<=0] = 0
pred_red[pred_red>=10] = 10

rmse(reduced_model$fitted.values, train_df$Q49)
rmse(pred_red[,], test_df$Q49)
plot(test_df$Q49, pred_red[,], ylab="Predicted Life Satisfaction", xlab="Recorded Life Satisfaction")
abline(0,1, col="red")
abline(1.5,1, col="blue")
abline(-1.5,1, col="blue")
```



## plot interactions and main effects
```{r}
plot_df = train_df
plot_df$y=reduced_model$fitted.values


# which(coef(summary(reduced_model))[, "Pr(>|t|)"] < 0.05)
# which(coef(summary(new_model))[, "Pr(>|t|)"] < 0.05)

# interaction term Q275 Q281
ggplot(plot_df, aes(x=Q275, y=y, fill=Q281)) + geom_boxplot() + scale_x_discrete(guide = guide_axis(angle = 40)) + facet_wrap(Q281 ~ .)

# interaction term Q276 Q282
ggplot(plot_df, aes(x=Q276, y=y, fill=Q282)) + geom_boxplot() + scale_x_discrete(guide = guide_axis(angle = 40)) + facet_wrap(Q282 ~ .)

# interaction term Q282 Q286
ggplot(plot_df, aes(x=Q282, y=y, fill=Q286)) + geom_boxplot() + scale_x_discrete(guide = guide_axis(angle = 40)) + facet_wrap(Q286 ~ .)
```



